---
title: "Part-3"
author: "Aylin Kallmayer"
date: "7/7/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
if (!require("devtools")) {
  install.packages("devtools", dependencies = TRUE)}
devtools::install_github("SebastianKorinth/eyEdu") 
library(eyEdu)
```

We will now look into eye tracking data recorded while participants had to read sentences on a screen. Each sentence was 1 trial. We will use the "eyEdu"-package because it is great for making your first steps in ET, also sebastian and I wrote it... :D check out more here: https://github.com/SebastianKorinth/eyEdu. The experiment was recorded with a cheap 60 Hz stationary eye tracker.

### Getting the reading experiment example data:  
```{r}
EyEduGetExamples(experiment.type = "reading")
```

For some eyEdu functions we have to define the parameter "raw.data.path", which is the path, where raw data were stored. For the example reading experiment (downloaded into your current working directory before) you can define this with:  
```{r}
raw.data.path <- paste(getwd(),"/exampleDataReadingExperiment-master/", sep = "")
```

### Importing raw eye movement data into eyEdu format  
EyEduImportRawData() reads raw data (.tsv files), separates eye movement from message information, extracts relevant information (e.g., screen dimensions, sample rate etc.), and saves these data for all participants into one structured file called eyEdu\_data.Rda within the raw data folder.
In this example, eight participants read ten single line sentences and responded afterwards to comprehension questions. Since we are not interested in eye movements associated with question processing, we will segment each trial into periods of interest (poi), that are limited by the messages "start trail" (= default, not explicitly parsed to the function) until "response\_time\_key\_finish\_reading".  
```{r}
EyEduImportEyetribeData(poi.end = "response_time_key_finish_reading")
```

### Visualizing single trials  
Now that we have our raw data in the eyEdu format, we can plot the raw data of single trials using the function EyEduPlotTrial().
The minimum requirement for this function is that we have to define for which participant (either name or number) and which trial number we would like to get a plot. There are several other parameters we can set for the function (e.g., color of samples for the left and right eye etc.). Note, the function will always try to superimpose eye movement data onto a screenshot of the stimulus participants saw during that time.  
```{r, warning=FALSE}
EyEduPlotTrial(participant.name = "Rosa", 
               trial.nr = 1)
```

### Low-pass filter  
In order to reduce "high frequency noise" - funny expression for EyeTribe data recorded at a 60 Hz sample rate - 
you might want to apply a low pass filter using the function EyEduLowPassFilter(). The default setting is a moving 
average over three samples (filter.settings = rep(1/3, 3)), which however we can be adjusted if necessary.  
```{r}
EyEduLowPassFilter()
```

Compare the plot above for the same participant and trial, this time plotting the filtered data (i.e., parameter show.filtered set to TRUE); data look "smoother".  
```{r, warning=FALSE}
EyEduPlotTrial(participant.name = "Rosa", 
               trial.nr = 1,
               show.filtered = TRUE)
```

### Fixation detection  
The function EyEduDetectFixationsIDT() uses the emov library by Simon Schwab and its I-DT algorithm 
(Salvucci & Goldberg, 2000), which uses dispersion limits at a minimum time periode to distinguish between 
fixations and saccades. We can parse the option to run fixation detection on raw or on filtered data.  
```{r}
EyEduDetectFixationsIDT(dispersion.var = 90,duration.var = 6,use.filtered = F)
```

Using the exact same code as above, we will have automatically added fixations to our trial plot.  
```{r, warning=FALSE}
EyEduPlotTrial(participant.name = "Rosa", 
               trial.nr = 1)
```

### Diagnostic plots  
A convenient way to visually inspect whether fixation detection led to meaningful results is - at least for reading experiments, in which the primary movement direction would be on a horizontal line - to plot the x-position of raw data samples (black line) over a short period of time (e.g., 2500 ms) superimposed by grey shaded areas indicating periods defined as fixations. In the example below we can see that fixation detection worked quite well:  
```{r}
EyEduPlotDiagnostics(participant.name = "Vladimir",
                     trial.nr = 5,sample.length = 2500, show.filtered = F)
```

Here an example in which fixation detection was less successful.  
```{r, warning=FALSE}
EyEduPlotDiagnostics(participant.name = "Clara",
                     trial.nr = 3,
                     sample.length = 2500, 
                     show.filtered = F)
```

### Areas of interest (aoi) around single words  
The following is meaningful for reading experiments only:  
The EyEduDefineWordAois() function parses screenshots made during the reading experiments defines areas of interest around single words.
Several parameters can be adjusted here to account for differences such as font size.  
```{r}
EyEduDefineWordAois(line.margin = 90,sparse.aoi.definition = T)
```

Using the EyEduPlotTrial() function, we will have automatically added the aoi info to our trial plot:  
```{r, warning=FALSE}
EyEduPlotTrial(participant.name = "Rosa", 
               trial.nr = 3,
               sparse.aoi.definition = T,
               aoi.names.screenshot = T)
```

Get labels for areas of interest:  
The OpenSesame experiment comprises messages about which sentence a participant read in which trial. 
The single words of each sentence can easily be allocated to each area of interest. 
An extra aoi is added in our case standing for a fixation point at the right midline of the screen, 
which participants had to fixate whenever they have finished reading a sentence.  
```{r}
EyEduImportWordAoiLabels(extra.aoi = c("fixation_point"), sparse.aoi.definition = T)
```

Assigning fixations to areas of interest:  
Now that we have prepared our areas of interest (including labels) we can assign each fixation to an aoi. 
Of course, some fixations will fall outside of aois.  
```{r}
EyEduAssignFixationsAois(sparse.aoi.definition = T,aoi.names.screenshot = T)
```

So we get the information, whether at all, for how long, and how often each participant fixated each word.  
```{r}
load("exampleDataReadingExperiment-master/eyEdu_data.Rda")
head(eyEdu.data$participants$Karl.tsv$fixation.data[,c(3,4,5,11)])
```

### Fixation summary  
The function EyEduGetFixationSummary() collects the fixation information of all participants and saves it 
into one data frame (fixation\_summary.Rda within the raw data folder). We can use this for further statistical 
analyses (e.g., testing whether longer words received more fixations etc.).  
```{r}
EyEduGetFixationSummary()
load("exampleDataReadingExperiment-master/fixation_summary.Rda")
fixation.summary[100:109,]
```

Using the fixation.summary data frame and your new aggregating and plotting skills, you can now plot and analyze summary statistics about specific Areas of Interest. But this might be more fun to do with the other data set you downloaded - the visual search data set.
So go on and reapply the steps you have just learned to search (see part 4)

