---
title: "Part-5"
author: "Aylin Kallmayer"
date: "7/8/2020"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(echo = TRUE)
if (!require("devtools")) {
  install.packages("devtools", dependencies = TRUE)}
devtools::install_github("SebastianKorinth/eyEdu") 
library(eyEdu)
library(dplyr)
library(Rmisc)
library(ggplot2)
# define screen size the experiment ran on for the plots to make sense
height <- 1050
width <- 1680
```

Now that you have the summary data frame for the search experiment, we will want to add the behavioral data to that data frame. This will allow us to conduct the analysis we need.

Now we will import the behavioral data (the .csv files in our exampleDataSearchExperiment-master folder). We can select the variables we want  
```{r}
raw.data.path <- paste(getwd(),"/part2/exampleDataSearchExperiment-master/", sep = "")
# behav_dat<-EyEduImportBehavioralData(raw.data.path,
#                                     selection = c("subject_nr", "count_trial_sequence",
#                                                    "correct", 
#                                                    "response_time", 
#                                                    "gaze_response",
#                                                    "stimulus", 
#                                                    "condition"))
files <- list.files(raw.data.path , pattern = "csv", recursive = F)

behav_dat <- NULL
for(file in files){
  print(paste(raw.data.path, file, sep = ""))
  temp <- read.csv(paste(raw.data.path, file, sep = ""),sep = ",", header = T, stringsAsFactors = F)
  behav_dat <- rbind(behav_dat,temp) 
}

behav_dat <- select(behav_dat, subject_nr, count_trial_sequence, correct, response_time, gaze_response, stimulus, condition)
```

### Behavioral analysis  
1) Get a "sense" of your data frame  
2) access individial variables  
3) choose only a subsection of your data  

Let's start with just the behavioral data. This is currently stored in a data frame called behav_dat  
```{r}
load("./part2/exampleDataSearchExperiment-master/fixation_summary.Rda")
str(behav_dat)
summary(behav_dat)
summary(fixation.summary)
str(fixation.summary)
# convert the "correct" variable to numeric
behav_dat$correct <- as.numeric(behav_dat$correct)
```

Get the mean response_time for each condition:  
```{r}
(agg <- aggregate(response_time ~ condition, mean, data=behav_dat))
```

Plot the RTs split by condition:  
```{r}
ggplot(agg, aes(x=condition, y=response_time, fill=condition))+
  geom_bar(stat="identity")+
  labs(x="Condition", y="Response time", title="Response-times over conditions")+
  scale_fill_manual(values=c("#cfcfea", "#732c2c"))+
  theme_classic()
```

Merge the behavioral data with the fixation data and clean:    
```{r}
clean_data <- fixation.summary %>%
  dplyr::rename(subject_nr = participant.nr, 
         count_trial_sequence = trial.index) %>%
  full_join(behav_dat) %>%
  filter(response_time > 100 & response_time < 5000, # clean data - take out rows where RTs were lower than 100 and more than 5000
         correct == 1 & !is.na(correct) & !is.na(aoi.label)) # also take only the correct trials
  
```

### Plot fixations  
We are again using ggplot, but this time we are using something called geom_point, seperated by condition:  
```{r, message=FALSE}
ggplot(clean_data, aes(x=fix.pos.x, y=fix.pos.y, color=condition))+
  geom_point(size=1)+
  coord_cartesian(xlim=c(0, width), ylim=c(0,height))+
  scale_y_reverse()+
  labs(x="Fixation position x", y="Fixation position y", title="Fixation positions")+
  scale_color_manual(values=c("#e7bb41", "#533a71"))+
  theme_classic()
```

Take duration into account as a weight for the circle size/colour/etc. Make sure the constant you pick works with your fixation duration values
try adjusting the constant and see what happens to your plot.  
```{r}
(clean_data$weight <- clean_data$fix.duration/10)
(clean_data$size <- clean_data$fix.duration/100)

ggplot(clean_data, aes(x=fix.pos.x, y=fix.pos.y, color=condition))+
  geom_point(aes(alpha=weight, size=size))+
  facet_grid(~condition)+
  coord_cartesian(xlim=c(0, width), ylim=c(0,height))+
  scale_y_reverse()+
  guides(alpha=F, size=F)+
  labs(x="Fixation position x",y="Fixation position y", title="Fixation positions over conditions")+
  scale_color_manual(values=c("#e7bb41", "#533a71"))+
  theme_classic()
```

OK, this was fun, but now we should actually try to see if participants looked longer at the inconsistent objects. We already know that they were slower, remember?...
In our data frame  we actually have marked the fixation which belong to the Interest Area of the object.
Now you can aggregate these fixations and check if the dwell time in the interest area differes between condition, so we could check if the Decision time was affected - a sub-part of the RTs.  
```{r}
agg_dwell <- aggregate(fix.duration ~ subject_nr+count_trial_sequence+condition+aoi.label, sum, data=clean_data)
(agg <- aggregate(fix.duration ~ condition + subject_nr, mean, data=agg_dwell))
```

Now plotting the results:  
```{r}
table <- summarySEwithin(withinvars = "condition", idvar = "subject_nr", measurevar = "fix.duration"
                         , data = agg)
ggplot(table, aes(x=condition, y=fix.duration, fill=condition))+
  geom_bar(stat="identity")+
  geom_errorbar(aes(ymin=fix.duration-sd, ymax=fix.duration+sd), width=.3)+
  labs(y="Dwell times", x="Condition", title="Dwell times over condition")+
  scale_fill_manual(values=c("#cfcfea", "#732c2c"))+
  theme_classic()
```





