---
title: "Part-4"
author: "Aylin Kallmayer"
date: "7/8/2020"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(echo = TRUE)
if (!require("devtools")) {
  install.packages("devtools", dependencies = TRUE)}
devtools::install_github("SebastianKorinth/eyEdu") 
library(eyEdu)
```

Applying the steps learnt in part three on the search experiment:  
```{r}
raw.data.path <- paste(getwd(),"/exampleDataSearchExperiment-master/", sep = "")
EyEduImportEyetribeData()
```

Plotting a single trial:  
```{r, warning=FALSE}
EyEduPlotTrial(participant.nr = 1, 
               trial.nr = 1)
```

Applying a filter and plotting the same trial with filtered data:  
```{r, warning=FALSE}
EyEduLowPassFilter()
EyEduPlotTrial(participant.nr = 1, 
               trial.nr = 1,
               show.filtered = TRUE)
```

Looks much less squiggly now :P

Detect fixations:  
Now plotting a different trial to keep things from geting boring.  
```{r, warning=FALSE}
EyEduDetectFixationsIDT(dispersion.var = 90,duration.var = 6,use.filtered = F)
EyEduPlotTrial(participant.nr = 1, 
               trial.nr = 3)
```

Now, checking whether fixation detection was succesful on different trials:  
```{r, warning=FALSE}
EyEduPlotDiagnostics(participant.nr = 1,
                     trial.nr = 1,sample.length = 2500, show.filtered = F)
EyEduPlotDiagnostics(participant.nr = 1,
                     trial.nr = 2,sample.length = 2500, show.filtered = F)
EyEduPlotDiagnostics(participant.nr = 1,
                     trial.nr = 3,sample.length = 2500, show.filtered = F)
```

Visually, it seems like the detection didnt work quite as well for this participant. Lets check out another participant:  
```{r}
EyEduPlotDiagnostics(participant.nr = 4,
                     trial.nr = 1,sample.length = 2500, show.filtered = F)
```

Not great either, but maybe a bit better than for participant 1?

Plotting AOIs:  
```{r}
EyEduImportAoIs(append.aois = FALSE,
                delete.existing.aois = TRUE ,
                screen.align.x = 1024,
                screen.align.y = 768)

EyEduAssignFixationsAois()
```

```{r, warning=FALSE}
EyEduPlotTrial(participant.nr = 1, 
               trial.nr = 3,
               fix.color = "green",
               show.filtered = FALSE,
               aoi.color = "black")
```

Get fixation summary:  
```{r}
EyEduGetFixationSummary()
load("exampleDataSearchExperiment-master/fixation_summary.Rda")
fixation.summary[100:109,]
```

Note: somehow drawing AOIs doesnt work and the assigned AOIs dont show up in the fixation summary ... :/ 